Tutorial_2a

### 2024-11-24 Tutorial_2a.Rmd
update tuts

Tutorial 2a shows various ways to fit Bayesian binary regression models to discrete time-to-event data. The data are taken from the first experiment of Panis and Schmidt (2016), but using only the no-mask trials (prime = blank, congruent, or incongruent). We perform prior predictive checks to infer the prior distributions we will employ. We compare models using WAIC and LOO, and interpret parameter estimates for one model. 
We end by plotting the logit and cloglog link functions, and illustrate how various prior distributions on the logit and cloglog scales look on the original probability (i.e., hazard) scale.

@@

```{r load-data}
ptb_data <- read_csv("Tutorial_1_descriptive_stats/data/inputfile_hazard_modeling.csv")
print(ptb_data,n=30)
summary(ptb_data) # 26602 rows: 6 participants, trial (experiment-wise), 3 conditions, 15 periods, and event indicator (0/1)
```

@@

# create categorical predictor for TIME named "timebin" with index coding
  mutate(timebin = factor(period,levels=c(6:15)),
# create continuous predictor for TIME named "period_9", centered on bin 9,
         period_9 = period - 9,
# create binary variables "dummys" to indicate each bin in the analysis
         d6  = if_else(period == 6, 1, 0),
         d7  = if_else(period == 7, 1, 0),
         d8  = if_else(period == 8, 1, 0),
         d9  = if_else(period == 9, 1, 0),
         d10 = if_else(period == 10, 1, 0),
         d11 = if_else(period == 11, 1, 0),
         d12 = if_else(period == 12, 1, 0),
         d13 = if_else(period == 13, 1, 0),
         d14 = if_else(period == 14, 1, 0),
         d15 = if_else(period == 15, 1, 0),

# create continuous predictor for trial number named "trial_c", centered on bin 1000 and rescaled
         trial_c = (trial - 1000)/1000,
# create categorical predictor for trial number named "stage" (early,middle,late) with index coding
         stage = ifelse(trial <= 500, 1, ifelse(trial > 1000, 3, 2)),
         stage = factor(stage, levels=c(1,2,3)),

@@

head(ptb_data,n=17)
summary(ptb_data)
```

@@

The first model we fit is a "random intercepts" model, where we fit a single intercept for each timebin and allow each intercept to vary between participants.
This is a general specification of TIME, and can be used if we do not want to make assumptions about how cloglog-hazard changes over time (within a trial). 
There are two ways to implement such a model. First, we can use the index-coding approach, which supplies an intercept for each level of TIME.

```{r data-priors-M0}
data_M0i <- ptb_data %>% select(pid, event, timebin)

@@

Perform a prior predictive check.

@@

model_M0_prior <-                     # wrong data file name
   brm(data = M0_data,
       family = bernoulli(link="cloglog"),
       event ~ 0 + timebin + (0 + timebin | pid),
       prior = priors_M0,
       chains = 4, cores = 4, iter = 3000, warmup = 1000,
       control = list(adapt_delta = 0.999, step_size = 0.04, max_treedepth = 12),
       seed = 12, init = "0",
       sample_prior = 'only',
       file = "Tutorial_2_Bayesian/models/model_M0_prior")
```

```{r}

#pp_check(model_M0_prior, stat="mean")

#http://bruno.nicenboim.me/bayescogsci/

# prior predictive check for 1 time bin
cloglog_predictive_distribution <-
  function(samples, N_obs) {
    # empty data frame with headers:
    df_pred <- tibble(trial = numeric(0),
                      y_pred = numeric(0),
                      iter = numeric(0))
    # i iterates from 1 to the length of samples,
      for (i in seq_along(samples)) {
        cloglog_haz <- samples[i]
        df_pred <- bind_rows(df_pred,
                             tibble(trial = 1:N_obs,
                                    y_pred = rbernoulli(N_obs, 1-exp((-1)*exp(cloglog_haz))),
                                    iter = i))
    }
    df_pred # return
  }
# simulate prior hazard values for 1 bin
N_samples <- 3000
N_obs <- 100 #nrow(M0_data)
#cloglog_samples <- rskew_normal(N_samples,-0.59,1.26,-4.22) # flat prior for cloglog intercept
cloglog_samples <- rskew_normal(N_samples,-1,1,-2) # weakly informative prior for cloglog intercept

prior_pred <- cloglog_predictive_distribution(cloglog_samples,N_obs)

prior_pred_haz <- prior_pred %>% group_by(iter) %>% summarise(pred_haz = mean(y_pred))

# plot 
ggplot(prior_pred_haz, aes(x=pred_haz)) +
  geom_histogram(binwidth=.01) +
  scale_x_continuous(limits = c(0,1))
```

```{r}
# Generate prior predictive samples for 10 bins 
bins_cloglog_predictive_distribution <-
  function(samples_bin, N_obs, N_bins) {
    # empty data frame with headers:
    df_pred <- tibble(trial = numeric(0),
                      y_pred = numeric(0),
                      iter = numeric(0),
                      bin = numeric(0))
    # i iterates from 1 to the length of samples,
    # b interates from 1 to N_bins
    for( b in 1:N_bins){
      for (i in 1:dim(samples_bin)[2]) {
        cloglog_haz <- samples_bin[b,i]
        df_pred <- bind_rows(df_pred,
                             tibble(trial = 1:N_obs,
                                    y_pred = rbernoulli(N_obs, 1-exp((-1)*exp(cloglog_haz))),
                                    iter = i,
                                    bin = b))
      }
    }
    df_pred # return
  }

# simulate prior hazard functions for 10 bin
N_samples <- 40
N_obs <- 100
N_bins <- 10
#cloglog_samples <- rskew_normal(N_samples,-0.59,1.26,-4.22) # flat prior for cloglog intercept
cloglog_samples_bin <- data.frame()
for(b in 1:N_bins){
  cloglog_samples_bin[b,1:N_samples] <- rskew_normal(N_samples,-1,1,-2) # weakly informative prior for cloglog intercept
}

prior_pred <- bins_cloglog_predictive_distribution(cloglog_samples_bin,N_obs,N_bins)

prior_pred_haz <- prior_pred %>% group_by(iter,bin) %>% summarise(pred_haz = mean(y_pred))

# plot 
ggplot(prior_pred_haz, aes(x=bin, y=pred_haz, group=iter)) +
  #geom_histogram(binwidth=.01) +
  geom_line(color="black") +
  geom_line(data = prior_pred_haz %>% filter(iter==4), color="red", linewidth = 2) +
  geom_line(data = prior_pred_haz %>% filter(iter==18), color="blue", linewidth = 2) +

  scale_y_continuous(limits = c(0,1)) +
  scale_x_continuous(limits = c(1,10), breaks = c(1:10))

```

```{r}
# Generate prior predictive hazard functions for 6 participants
newdata_prior = tibble(timebin = c(6:15)) %>% 
                            expand_grid(pid = 1:6)
N_obs = 100

@@

for(i in 1:N_obs){
  prior_pred <- add_predicted_draws(model_M0_prior, newdata = newdata_prior, summary=F, ndraws=NULL)
  prior_pred <- prior_pred %>% mutate(obs = i)
  df_pred <- bind_rows(df_pred, prior_pred)
}
head(df_pred)

# calculate hazard per draw (average across obs)
df_pred_haz <- df_pred %>% group_by(pid,timebin,.draw) %>% summarise(pred_haz = mean(.prediction))
summary(df_pred_haz)

# plot 40 draws per participant
ggplot(df_pred_haz %>% filter(.draw < 21), aes(x=timebin, y=pred_haz, group=.draw)) +

   geom_line(color="black") +
   geom_line(data=df_pred_haz %>% filter(.draw == 20), color="red", linewidth=2) +

  scale_y_continuous(limits = c(0,1)) +
  scale_x_continuous(limits = c(6,15), breaks = c(6:15)) +
  facet_wrap(~pid)

```

Fit model M0i.

@@

```{r}
model_M0i <- readRDS("Tutorial_2_Bayesian/models/model_M0i.rds")
summary(model_M0i)
formula(model_M0i)
plot(model_M0i)
```

@@

```{r data-priors-M0d}
data_M0d <- ptb_data %>% select(pid, event, d6:d8, d10:d15)

@@

model_M0d <-      
   brm(data = data_M0d,
       family = bernoulli(link="cloglog"),
       event ~ 0 + d6 + d7 + d8 + Intercept + d10 + d11 + d12 + d13 + d14 + d15 + 
                   (d6 + d7 + d8 + 1 + d10 + d11 + d12 + d13 + d14 + d15  | pid),
       prior = priors_M0d,
       chains = 4, cores = 4, iter = 3000, warmup = 1000,
       control = list(adapt_delta = 0.999, step_size = 0.04, max_treedepth = 12),
       seed = 12, init = "0",
       file = "Tutorial_2_Bayesian/models/model_M0d")
```

@@

```{r}
model_M0d <- readRDS("Tutorial_2_Bayesian/models/model_M0d.rds")
formula(model_M0d)
fixef(model_M0d)
```

### Make assumptions to simplify the model

Instead of using a general specification of TIME, we can treat TIME as a continuous variable and make assumptions about how hazard changes over time within a trial. For example, we might assume that hazard changes in a linear + quadratic fashion over time bins within a trial.

@@

Fit the model with a continuous predictor (M0c).

@@

```{r}
model_M0c <- readRDS("Tutorial_2_Bayesian/models/model_M0c.rds")
summary(model_M0c)
formula(model_M0c)
```

## Intermezzo: Prior predictive checks

model_M0c_prior <-
   brm(data = M0c_data,
       family = binomial(link="cloglog"),
       event | trials(1) ~ 0 + Intercept + period_9 + period_9_sq + period_9_cu + 
                          (0 + Intercept + period_9 + period_9_sq + period_9_cu  | pid),
       prior = priors_M0c,
       chains = 4, cores = 4, iter = 3000, warmup = 1000,
       control = list(adapt_delta = 0.999, step_size = 0.04, max_treedepth = 12),
       seed = 12, init = "0",
       sample_prior = "only",
       file = "Tutorial_2_Bayesian/models/model_M0c_prior")
```

```{r}
# Generate prior predictive hazard functions for 6 participants
newdata_prior = tibble(period_9 = c(-3:6)) %>% 
  mutate(period_9_sq = period_9^2,
         period_9_cu = period_9^3) %>%
  expand_grid(pid = 1:6)

N_obs = 100

# prepare columns of data set
df_pred <- tibble(period_9 = integer(0),
                  period_9_sq = double(0),
                  period_9_cu = double(0),
                  pid = integer(0),
                  .row = integer(0),
                  .chain = integer(0),
                  .iteration = integer(0),
                  .draw = integer(0),
                  .prediction = integer(0),
                  obs = integer(0))

for(i in 1:N_obs){
  prior_pred <- add_predicted_draws(model_M0c_prior, newdata = newdata_prior, summary=F, ndraws=NULL)
  prior_pred <- prior_pred %>% mutate(obs = i)
  df_pred <- bind_rows(df_pred, prior_pred)
}
head(df_pred)

# calculate hazard per draw (average across obs)
df_pred_haz <- df_pred %>% group_by(pid,period_9,.draw) %>% summarise(pred_haz = mean(.prediction))
summary(df_pred_haz)

# plot 40 draws per participant
ggplot(df_pred_haz %>% filter(.draw < 2), aes(x=period_9, y=pred_haz, group=.draw)) +

   geom_line(color="black") +
   geom_line(data=df_pred_haz %>% filter(.draw == 17), color="red", linewidth=2) +

  scale_y_continuous(limits = c(0,1)) +
  scale_x_continuous(limits = c(-3,6), breaks = c(-3:6)) +
  facet_wrap(~pid)
```

##

```{r}
dat <- tibble(x = -3:6,
              x2 = x^2,
              x3 = x^3,
              cloglog1 = -1.8   + 0.3*x   - 0.1*x2   + 0.009*x3,
              cloglog2 = -1     + 0.4*x   - 0.1*x2   + 0.007*x3,
              cloglog3 = -3     - 0.1*x   + 0.12*x2  - 0.002*x3,
              cloglog4 = -0.40  + 0.2*x   - 0.04*x2  + 0.003*x3,
              haz1 = 1 - exp(-1*exp(cloglog1)),
              haz2 = 1 - exp(-1*exp(cloglog2)),
              haz3 = 1 - exp(-1*exp(cloglog3)),
              haz4 = 1 - exp(-1*exp(cloglog4)))

p1<-ggplot(dat, aes(x=x)) +  
      geom_line(aes(y=cloglog1), color="black") +
      geom_line(aes(y=cloglog2), color="green") +
      geom_line(aes(y=cloglog3), color="red") +
      geom_line(aes(y=cloglog4), color="blue") +
      scale_y_continuous(limits=c(-6,2)) +
      scale_x_continuous(breaks=c(-3:6))

p2<-ggplot(dat, aes(x=x))+
      geom_line(aes(y=haz1), color="black") + 
      geom_line(aes(y=haz2), color="green") +
      geom_line(aes(y=haz3), color="red") +
      geom_line(aes(y=haz4), color="blue") +
      scale_y_continuous(limits=c(0,1)) +
      scale_x_continuous(breaks=c(-3:6))

```{r}
tibble(cloglog = rnorm(1000, .1, .1)) %>%
  ggplot(aes(x=cloglog)) +
  geom_density()
```

```{r data-priors-M0c}
M0c_data <- ptb_data %>% select(pid, event, period_9) %>%
  mutate(period_9_sq = period_9^2)

priors_M0c_prior <- c(
  #set_prior("skew_normal(-0.2,0.71,-2.2)", class = "b"),      
 set_prior("skew_normal(-1,1,-2)", class = "b", coef = "Intercept"),

#set_prior("normal(0.5,.5)",class="b",coef="period_9"),
#set_prior("normal(-.1,.2)",class="b",coef="period_9_sq"),   # Np3

#set_prior("normal(0.1,.08)",class="b",coef="period_9"), # ok !!!
#set_prior("normal(0,.06)",class="b",coef="period_9_sq"),

set_prior("normal(0.1,.1)",class="b",coef="period_9"), # ok !!!
set_prior("normal(0,.06)",class="b",coef="period_9_sq"),
#set_prior("normal(0,.003)",class="b",coef="period_9_cu"),

#set_prior("skew_normal(0.3,0.1,2)",class="b",coef="period_9"), #skNp3
#set_prior("skew_normal(0,0.1,0)",class="b",coef="period_9_sq"),

 set_prior("normal(0, .01)", class = "sd"),  
  set_prior("lkj(2)", class = "cor")  
)
```

model_M0_lin_prior454 <-
   brm(data = M0c_data,
       family = binomial(link="cloglog"),
       event | trials(1) ~ 0 + Intercept + period_9 + period_9_sq  +
                          (0 + Intercept + period_9 + period_9_sq  | pid),
       prior = priors_M0c_prior,
       chains = 4, cores = 4, iter = 3000, warmup = 1000,
       control = list(adapt_delta = 0.999, step_size = 0.04, max_treedepth = 12),
       seed = 12, init = "0",
       sample_prior = "only",
       file = "Tutorial_2_Bayesian/models/model_M0_lin_prior454")
```

```{r}
#model_M0_lin_prior448 <- readRDS("Tutorial_2_Bayesian/models/model_M0_lin_prior448.rds")

post <- as_draws_df(model_M0_lin_prior454) %>% 
  select(-lp__) %>% 
  as_tibble()

tidy_fixed <- post %>% 
  select(starts_with("b_"), .chain, .iteration, .draw) %>% 
  rename(chain=.chain, iter=.iteration, draw=.draw) %>% 
  pivot_longer(-c(chain, draw, iter)) 
head(tidy_fixed)
tail(tidy_fixed)

# plot
p_tidy_fixed <- ggplot(tidy_fixed, aes(x = name, y = value, fill = name)) +  
  stat_halfeye(alpha=0.7, point_interval = "median_qi",.width=c(.8,.99) ) 
p_tidy_fixed

```

```{r}
#model_M0_lin_prior14 <-readRDS("Tutorial_2_Bayesian/models/model_M0c_lin_prior14.rds")               
# Generate prior predictive hazard functions for 6 participants
newdata_prior = tibble(period_9 = c(-3:6)) %>% 
  mutate(period_9_sq = period_9^2) %>%
  expand_grid(pid = 1:6)

N_obs = 100 # increase for smoother results

# prepare columns of data set
df_pred <- tibble(period_9 = integer(0),
                 period_9_sq = integer(0),
                # period_9_ce = integer(0),
                 pid = integer(0),
                  .row = integer(0),
                  .chain = integer(0),
                  .iteration = integer(0),
                  .draw = integer(0),
                  .prediction = integer(0),
                  obs = integer(0))

set.seed(1)

for(i in 1:N_obs){
  prior_pred <- add_predicted_draws(model_M0_lin_prior454, 
                                    newdata = newdata_prior, 
                                    summary=F, 
                                    ndraws=NULL) %>% 
                mutate(obs = i)
  df_pred <- bind_rows(df_pred, prior_pred)
}

head(df_pred)

# calculate hazard per draw (average across obs)
df_pred_haz <- df_pred %>% 
  group_by(pid,period_9,.draw) %>% 
  summarise(pred_haz = mean(.prediction)) 
summary(df_pred_haz)

# plot 40 draws per participant
sel = sample(1:8000,3,replace=F)

ggplot(df_pred_haz %>% 
   filter(.draw <639), aes(x=period_9, y=pred_haz, group=.draw)) +
   geom_line(color="black") +
   geom_line(data=df_pred_haz %>% filter(.draw ==sel[1]), 
             color="red", linewidth=2) +
  geom_line(data=df_pred_haz %>% filter(.draw == sel[2]), 
            color="green", linewidth=2) +
  geom_line(data=df_pred_haz %>% filter(.draw == sel[3]), 
          color="blue", linewidth=2) +
  scale_y_continuous(limits = c(0,1)) +
  scale_x_continuous(limits = c(-3,6), breaks = c(-3:6)) +
  facet_wrap(~pid)
```

@@

The next model we fit also includes our predictor variable prime type.
First, we can use index coding for the categorical predictor.

@@

```{r data-priors-M1}
M1_data <- ptb_data %>% select(pid, event, timebin, prime)

priors_M1 <- c(
  set_prior("skew_normal(-0.59,1.26,-4.22)", class = "b"), 
  set_prior("normal(0, 1)", class = "sd"),                    
  set_prior("lkj(2)", class = "cor")                        
)
```

Fit model M1.

```{r fit-model-M1, eval=F}
plan(multicore)

model_M1 <-
   brm(data = M1_data,
       family = binomial(link="cloglog"),
       event|trials(1) ~ 0 + timebin:prime  +
                        (0 + timebin:prime | pid),
       prior = priors_M1,
       chains = 4, cores = 4, iter = 3000, warmup = 1000,
       control = list(adapt_delta = 0.999, step_size = 0.04, max_treedepth = 12),
       seed = 12, init = "0",
       file = "Tutorial_2_Bayesian/models/model_M1")
```

Model M1 took about 126 minutes to fit.

```{r}
model_M1 <- readRDS("Tutorial_2_Bayesian/models/model_M1.rds")
summary(model_M1)
formula(model_M1)
```

Second, if we want to make assumptions about (1) how hazard changes over TIME in the reference condition (blank prime), and (2) how the effect of congruent and incongruent primes change over TIME (relax the proportionality assumption), then we can switch to a dummy coding approach and treat TIME as a continuous variable.
For example, we may assume that hazard can change in a linear + quadratic + cubic fashion over time for a blank prime, and that the effects of congruent and incongruent primes relative to blank change in a linear + quadratic + cubic fashion, and fit the following model (M1d).

```{r data-priors-M1d}
M1d_data <- ptb_data %>% select(pid, event, period_9, condition) %>%

>>>>

Tutorial 2a shows various ways to fit Bayesian binary regression models to discrete time-to-event data. The data are taken from the first experiment of Panis and Schmidt (2016), but using only the no-mask trials (with factor prime = blank, congruent, or incongruent). We perform some prior predictive checks to infer the prior distributions we will employ. We compare models using WAIC and LOO, and interpret parameter estimates for one model. 
We end by plotting the logit and cloglog link functions, and illustrate how various prior distributions on the logit and cloglog scales look on the original probability (i.e., hazard) scale.

@@

```{r load-data}
ptb_data <- read_csv("Tutorial_1_descriptive_stats/data/inputfile_hazard_modeling.csv")
head(ptb_data)
```

@@

# create categorical predictor for TIME named "timebin" with index coding
  mutate(timebin = factor(period,levels=c(6:15)),
# create continuous predictor for TIME named "period_9", centered on bin 9,
         period_9 = period - 9,
# create binary variables to indicate each bin
         d6  = if_else(period == 6, 1, 0),
         d7  = if_else(period == 7, 1, 0),
         d8  = if_else(period == 8, 1, 0),
         d9  = if_else(period == 9, 1, 0),
         d10 = if_else(period == 10, 1, 0),
         d11 = if_else(period == 11, 1, 0),
         d12 = if_else(period == 12, 1, 0),
         d13 = if_else(period == 13, 1, 0),
         d14 = if_else(period == 14, 1, 0),
         d15 = if_else(period == 15, 1, 0),

# create continuous predictor for trial number named "trial_c", centered on bin 1000 and rescale
         trial_c = (trial - 1000)/1000,
# create categorical predictor for trial number named "stage" (early,middle,late) with index coding
         stage = ifelse(trial <= 500, 1, ifelse(trial > 1000, 3, 2)),
         stage = factor(stage, levels=c(1,2,3)),

@@

head(ptb_data,n=17)
```

@@

The first model we fit is a "random intercepts" model, where we fit a single grand intercept for each timebin and add random intercepts that vary between participants.
This constitutes a general specification of TIME, and can be used if we do not want to make assumptions about how cloglog-hazard changes over time (within a trial).

There are two ways to implement such a model. First, we can use the index-coding approach, which provides an intercept for each level of TIME (variable timebin in ptb_data).

Prepare the data file, and specify priors. The skew_normal prior is set for each grand intercept on the cloglog scale, and should reflect our prior beliefs.

```{r data-priors-M0i}
data_M0i <- ptb_data %>% select(pid, event, timebin)

@@

Perform a prior predictive check to see if the binary observations generated from the sprecified prior distributions reflects our prior beliefs. First, sample the prior distributions using sample_prior="only".

@@

model_M0i_prior <-                  
   brm(data = data_M0i,
       family = bernoulli(link="cloglog"),
       formula = event ~ 0 + timebin + (0 + timebin | pid),
       prior = priors_M0i,
       chains = 4, cores = 4, 
       iter = 3000, warmup = 1000,
       control = list(adapt_delta = 0.999, 
                      step_size = 0.04, 
                      max_treedepth = 12),
       seed = 12, init = "0",
       sample_prior = "only",
       file = "Tutorial_2_Bayesian/models/model_M0i_prior")
```

Next, use them to predict prior data. To better understand what the function add_predicted_draws is doing (which we will typically use), we first simulate prior data for one timebin manually.

```{r ppc-manual}
# prior predictive check for 1 time bin
# see http://bruno.nicenboim.me/bayescogsci/

# cloglog and inverse-cloglog link functions
inverse_cloglog <- function(x) {return(1-(exp(-1*exp(x))))}
cloglog <- function(x) {return(log(-1*log(1-x)))}

# Specify
N_samples <- 1000 # the number of samples from the skew_normal
N_obs <- 100      # the number of simulated observations per sample
set.seed(1)

# function to generate binary observations from samples
cloglog_predictive_distribution <- function(samples, Nobs) {

    # empty data frame with headers
    df_pred <- tibble(trial = numeric(0),
                      y_pred = numeric(0),
                      iter = numeric(0))
    # i iterates from 1 to the length of samples
    for (i in seq_along(samples)) {
      cloglog_haz <- samples[i]
      df_pred <- bind_rows(df_pred,
          tibble(trial = 1:N_obs,
                 y_pred = rbinom(Nobs, 1, inverse_cloglog(cloglog_haz)),
                 iter = i))
    }
    df_pred 
 }

# sample prior cloglog-hazard values for 1 bin
cloglog_samples <- rskew_normal(N_samples,-1,1,-2)

# apply the function
prior_pred <- cloglog_predictive_distribution(cloglog_samples,N_obs)

# calculate hazard as the mean over binary observations
prior_pred_haz <- prior_pred %>% 
  group_by(iter) %>% 
  summarise(pred_haz = mean(y_pred))

# plot 
ggplot(prior_pred_haz, aes(x=pred_haz)) +
  geom_histogram(binwidth=.01) +
  scale_x_continuous(limits = c(0,1)) +
  labs(x = "simulated hazard",
       title = "Prior predictive distribution",
       subtitle = str_c(N_samples," samples; ",N_obs," observations per sample"))
```

The shape of the prior predictive distribution looks as expected (compare with Figure X in the supplementary material): hazard values below .5 are more likely than hazard values above .5.

Now use add_predicted_draws() to simulate prior-based data for 10 bins and 6 subjects, while take into account samples from the priors for the standard deviation of the random effects, and correlations between parameters.

```{r ppc-M0i-prior}
# Generate prior predictive hazard functions for 6 participants
newdata_prior = tibble(timebin = 6:15) %>% 
           expand_grid(pid = 1:6)

# Specify
N_obs = 100
set.seed(2)

@@

# Call add_predicted_draws() for each simulated observation
for(i in 1:N_obs){
  prior_pred <- add_predicted_draws(model_M0i_prior, 
                                    newdata=newdata_prior, 
                                    summary=F, 
                                    ndraws=NULL) %>% 
                mutate(obs = i)
  df_pred <- bind_rows(df_pred, prior_pred)
}

# calculate hazard per draw (average across observations)
df_pred_haz <- df_pred %>% 
  group_by(pid,timebin,.draw) %>% 
  summarise(pred_haz = mean(.prediction))

# plot 40 draws per participant
ggplot(data = df_pred_haz %>% filter(.draw < 41), 
       aes(x=timebin, y=pred_haz, group=.draw)) +
   geom_line(color="black") +
   geom_line(data=df_pred_haz %>% filter(.draw == 20), 
             color="red", 
             linewidth=2) +
  scale_y_continuous(limits = c(0,1)) +
  scale_x_continuous(limits = c(6,15), breaks = c(6:15)) +
  labs(x = "time bin",
       y = "simulated hazard",
       title = "Prior predictive distributions",
       subtitle = str_c("8000 samples; ",N_obs," observations per sample")) +
  facet_wrap(~pid)
```

Now we can fit model M0i to get the posterior distributions.

@@

```{r insppect-M0i}
model_M0i <- readRDS("Tutorial_2_Bayesian/models/model_M0i.rds")
summary(model_M0i)
formula(model_M0i)
plot(model_M0i)
```

@@

```{r data-priors-M0d}
# select bin indicator variables
data_M0d <- ptb_data %>% select(pid, event, d6:d8, d10:d15)

@@

model_M0d <-      
   brm(data = data_M0d,
       family = bernoulli(link="cloglog"),
       formula = event ~ 0 + d6 + d7 + d8 + Intercept + d10 + d11 + d12 + d13 + d14 + d15 + 
                   (d6 + d7 + d8 + 1 + d10 + d11 + d12 + d13 + d14 + d15  | pid),
       prior = priors_M0d,
       chains = 4, cores = 4, 
       iter = 3000, warmup = 1000,
       control = list(adapt_delta = 0.999, 
                      step_size = 0.04, 
                      max_treedepth = 12),
       seed = 12, init = "0",
       file = "Tutorial_2_Bayesian/models/model_M0d")
```

@@

```{r inpect-M0d}
model_M0d <- readRDS("Tutorial_2_Bayesian/models/model_M0d.rds")
formula(model_M0d)
fixef(model_M0d)
```

Third, we can make assumptions to simplify the model. Instead of using a general specification of TIME (bin rank), we can treat TIME as a continuous variable and make assumptions about how cloglog-hazard changes over time within a trial. For example, we might assume that cloglog-hazard changes in a linear ("period_9") + quadratic ("period_9_sq") fashion over time bins within a trial.

@@

Perform a prior predictive check to see if the binary observations generated from the sprecified prior distributions reflect our prior beliefs. 
The following code allows to change the coefficients and generate (cloglog-)hazard functions.
Once you have found the coefficients for period_9 and period_9_sq that generate a variety of hazard functions spanning your prior beliefs, you can then use them to estimate a mean and standard deviation for each coefficient.

```{r search-coef}
dat <- tibble(x = -3:6,
              x2 = x^2,
              cloglog1 = -1.8   + 0.3*x   - 0.1*x2,
              cloglog2 = -1     + 0.4*x   - 0.1*x2,
              cloglog3 = -3     - 0.1*x   + 0.12*x2,
              cloglog4 = -0.40  + 0.2*x   - 0.04*x2,
              haz1 = 1 - exp(-1*exp(cloglog1)),
              haz2 = 1 - exp(-1*exp(cloglog2)),
              haz3 = 1 - exp(-1*exp(cloglog3)),
              haz4 = 1 - exp(-1*exp(cloglog4)))

p1<-ggplot(dat, aes(x=x)) +  
      geom_line(aes(y=cloglog1), color="black") +
      geom_line(aes(y=cloglog2), color="green") +
      geom_line(aes(y=cloglog3), color="red") +
      geom_line(aes(y=cloglog4), color="blue") +
      scale_y_continuous(limits=c(-6,2)) +
      scale_x_continuous(breaks=c(-3:6)) +
      labs(x = "timebin", y = "cloglog-hazard")

p2<-ggplot(dat, aes(x=x))+
      geom_line(aes(y=haz1), color="black") + 
      geom_line(aes(y=haz2), color="green") +
      geom_line(aes(y=haz3), color="red") +
      geom_line(aes(y=haz4), color="blue") +
      scale_y_continuous(limits=c(0,1)) +
      scale_x_continuous(breaks=c(-3:6)) +
      labs(x = "timebin", y = "hazard")

First, sample the prior distributions, and set the variance of the normal prior for class "sd" to a low value, to minimize the effect of the random variables while inspecting the relation between the selected priors for period_9 and period_9_sq and the shape of the prior predictive distributions.

```{r priors-M0c-ppc}
priors_M0c_ppc <- c(
  set_prior("skew_normal(-1,1,-2)", class = "b", coef = "Intercept"),  
  set_prior("normal(0.1,.1)", class = "b", coef = "period_9"), 
  set_prior("normal(0,.06)", class= "b", coef="period_9_sq"),   
  set_prior("normal(0, .1)", class = "sd"),  # set standard deviation to a small value for now
  set_prior("lkj(2)", class = "cor")  
)
```

```{r fit-model-M0c-prior, eval=F}
plan(multicore)

model_M0c_prior <-
   brm(data = data_M0c,
       family = bernoulli(link="cloglog"),
       formula = event ~ 0 + Intercept + period_9 + period_9_sq +  
                          (0 + Intercept + period_9 + period_9_sq | pid),
       prior = priors_M0c_ppc,
       chains = 4, cores = 4, 
       iter = 3000, warmup = 1000,
       control = list(adapt_delta = 0.999, 
                      step_size = 0.04, 
                      max_treedepth = 12),
       seed = 12, init = "0",
       sample_prior = "only",
       file = "Tutorial_2_Bayesian/models/model_M0c_prior")
```

Check the prior distributions.

```{r prior-distr}
# extract prior draws
post <- as_draws_df(model_M0c_prior) %>% 
  select(-lp__) %>% 
  as_tibble()

tidy_prior <- post %>% 
  select(starts_with("b_"), .chain, .iteration, .draw) %>% 
  rename(chain=.chain, iter=.iteration, draw=.draw) %>% 
  pivot_longer(-c(chain, draw, iter))

# plot
ggplot(tidy_prior, aes(x = name, y = value, fill = name)) +  
  stat_halfeye(alpha=0.7, 
               point_interval = "median_qi",
               .width=c(.8,.99), 
               show.legend = F) +
  labs(x = "parameter",
       title = "Samples from the prior distributions")
```

Generate prior predictive check for each participant.

```{r ppc-model-M0c-prior}
# Generate prior predictive hazard functions for 6 participants
# Specify
N_obs = 100 # increase for smoother results
set.seed(1)

# make new data
newdata_prior = tibble(period_9 = c(-3:6)) %>% 
  mutate(period_9_sq = period_9^2) %>%
  expand_grid(pid = 1:6)

# prepare columns of data set with predictions
df_pred <- tibble(period_9 = integer(0),
                  period_9_sq = integer(0),
                  pid = integer(0),
                  .row = integer(0),
                  .chain = integer(0),
                  .iteration = integer(0),
                  .draw = integer(0),
                  .prediction = integer(0),
                  obs = integer(0))

# extract predictions
for(i in 1:N_obs){
  prior_pred <- add_predicted_draws(model_M0c_prior, 
                                    newdata = newdata_prior, 
                                    summary=F, 
                                    ndraws=NULL) %>% 
                mutate(obs = i)
  df_pred <- bind_rows(df_pred, prior_pred)
}

# calculate hazard per draw (average across obs)
df_pred_haz <- df_pred %>% 
  group_by(pid,period_9,.draw) %>% 
  summarise(pred_haz = mean(.prediction))

# plot 200 draws per participant, highlighting three
sel = sample(1:200,3,replace=F)

ggplot(df_pred_haz %>% 
   filter(.draw <200), aes(x=period_9, y=pred_haz, group=.draw)) +
   geom_line(color="black") +
   # highlight
   geom_line(data=df_pred_haz %>% filter(.draw ==sel[1]), 
             color="red", linewidth=2) +
   geom_line(data=df_pred_haz %>% filter(.draw == sel[2]), 
            color="green", linewidth=2) +
   geom_line(data=df_pred_haz %>% filter(.draw == sel[3]), 
          color="blue", linewidth=2) +
  scale_y_continuous(limits = c(0,1)) +
  scale_x_continuous(limits = c(-3,6), breaks = c(-3:6)) +
  facet_wrap(~pid)
```

A wide range of differently shaped hazard functions can be observed, and early bins are unlikely to have high hazard values, consistent with our prior beliefs. Note that some simulated hazard functions start off with a high hazard in bin -3. These reflect a combination of negative slopes and a positive coefficient for period_9_sq. These possibilities are required to encompass hazard functions that initially decrease and then increase, as observed in the empirical hazard functions for incongruent primes.

Now increase the variance again of the normal prior for class "sd", and fit model M0c to get the posterior distributions.

@@

```{r inspect-M0c}
model_M0c <- readRDS("Tutorial_2_Bayesian/models/model_M0c.rds")
summary(model_M0c)
formula(model_M0c)
```

@@

The next models we fit also include our predictor variable prime type.
First, we can use index coding for this categorical predictor.

@@

```{r data-priors-M1}
data_M1i <- ptb_data %>% select(pid, event, timebin, prime)

priors_M1i <- c(
  set_prior("skew_normal(-1,1,-2)", class = "b"), 
  set_prior("normal(0, 1)", class = "sd"),                    
  set_prior("lkj(2)", class = "cor")                        
)
```

Fit model M1i. The interaction between timebin and prime will specify 30 grand intercepts.

```{r fit-model-M1i, eval=F}
plan(multicore)

model_M1i <-
   brm(data = data_M1i,
       family = bernoulli(link="cloglog"),
       formula = event ~ 0 + timebin:prime  +
                        (0 + timebin:prime | pid),
       prior = priors_M1i,
       chains = 4, cores = 4, 
       iter = 3000, warmup = 1000,
       control = list(adapt_delta = 0.999, 
                      step_size = 0.04, 
                      max_treedepth = 12),
       seed = 12, init = "0",
       file = "Tutorial_2_Bayesian/models/model_M1i")
```

Model M1i took about 124 minutes to fit.

```{r inspect-M1i}
model_M1i <- readRDS("Tutorial_2_Bayesian/models/model_M1i.rds")
fixef(model_M1i)
formula(model_M1i)
```

Second, if we want to make assumptions about (1) how hazard changes over TIME in the reference condition (blank prime), and (2) how the effect of congruent and incongruent primes change over TIME (relax the proportionality assumption), then we can switch to a dummy coding approach and treat TIME as a continuous variable.
For example, we may assume that hazard can change in a linear + quadratic fashion over time for a blank prime, and that the effects of congruent and incongruent primes relative to blank change in a linear + quadratic fashion, and fit the following model (M1d).

```{r data-priors-M1d}
data_M1d <- ptb_data %>% select(pid, event, period_9, condition) %>%
  mutate(period_9_sq = period_9^2)
